---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  warning = F
)
```

# eccficm: Expected Conditional Characteristic Function-based Independence Criterion measures

<!-- badges: start -->
<!-- badges: end -->

The goal of `eccficm` is to provide new distance correlation/covariance type of measures for testing independence between two random vectors using the expected conditional characteristic function-based independence criterion (ECCFIC) method developed by @yin2019new and @ke2019expected. The method is nonparametric, can detect both linear and nonlinear dependencies, and overcomes many challenges/limitations of independence measures.
 

## Installation

You can install the released version of `eccficm` from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("eccficm")
```

## Available Functions

The main functions are `ecd`, `eccfic`, `ecd.test`, and `eccfic.test` as described below.
 
 - `ecd`, `eccfic`: compute the covariance and correlation type statistics.

 - `ecd.test`, `eccfic.test`: perform a permutation test of independence based on the ECD and ECCFIC correlation and covariance type statistics, respectively.

- `generateData`: generates optional data for feature screening based on
 some specific models.

 - `fscreen`: Performs the sure independence feature screening using
 correlation learning between each predictor and the response.

## Example

In the examples that follow, we use the aircraft data available in the `sm` package in R. The data contains six characteristics of aircraft designs in the twentieth century. We consider two variables, wing span (m) and maximum speed (km/h) in period 3 with 230 observations. Our goal is to test the independence of log(Span) and log(Speed) using ECCFIC correlations as test statistics. Results based on distance correlation is included for comparison. $999$ replicates are used for the permutation test. 


We first load the `sm` package to get access to the `aircraft` data and filter the `aircraft` data to obtain data for just the third brand time period.

```{r echo=T}
library(sm)

# get the aircraft data
data("aircraft", package = "sm")

# get data for the third brand time period
aircraft <- aircraft[aircraft$Period==3, ] 

attach(aircraft) # expose variable names to the R search path
```

Create a scatterplot to explore the relationship between `log(Speed)` and `log(Span)`.

```{r echo=T}
plot(log(Speed), log(Span))
```

The above graph suggests the existence of some sort of association between `Speed` and `Span` which appears to be nonlinear. In the subsequent codes we will use the hypothesis testing functions in the `eccficm` package as well as one from the `energy` package to determine if there is enough evidence to conclude on the observed relationship.


Here, we load our software package `eccficm`. 

```{r echo=T}
library(eccficm)
```

First, we test the independence of `log(Speed)` and `log(Span)` with the `ecd.test` function. We use 10 slices for the slicing estimation, ECD correlation as test statistic, and perform 999 permutations.

```{r echo=T}
ecd.test(log(Speed), log(Span), est="slice", ns=10, ts="ecdcor",  B=999)
```

Using the ECD correlation as test statistic, we obtain a p-value of **0.001**.

Similarly, we use the `eccfic.test` function with 10 slices for its slicing estimation, ECCFIC correlation as test statistic, and 999 total permutations. The p-value for this test is also **0.001**. 

```{r echo=T}
eccfic.test(log(Speed), log(Span), est="slice", ns=10, ts="eccficcor",  B=999)
```


Finally, we compare the results for our independence measures with that of the distance correlation with the same number of permutations. The DC independence testing function is avaialble in the `energy` package. Interestingly, using DC as test statistic produces the same p-value of **0.001**.

```{r echo=T}
energy::dcor.test(log(Speed), log(Span), R=999)
```


From the above outputs,  despite the nonlinear relation between wing span and maximum speed, the `ecd.test`, `eccfic.test`, and `dcor.test` functions returned  a very small p-value of 0.001 suggesting that there is enough evidence to conclude that wing span depends on maximum speed. This shows how sensitive distance correlation and ECCFIC correlations are to nonlinear associations. Moreover, we see how comparable our measures are to the distance correlation.

