---
output: github_document
bibliography: inst/REFERENCES.bib
link-citations: yes
linkcolor: blue
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  warning = F,
  eval = T
)
```

# eccficm: Expected Conditional Characteristic Function-based Independence Criterion Measures

<!-- badges: start -->
<!-- badges: end -->

The goal of `eccficm` is to provide new distance correlation/covariance type of measures for testing independence between two random vectors using the expected conditional characteristic function-based independence criterion (ECCFIC) method developed by @yin2019new and @ke2019expected. The method is nonparametric, can detect both linear and nonlinear dependencies, and overcomes many challenges/limitations of independence measures.

Although the package is based on both ECD measures [@yin2019new] and ECCFIC measures [@ke2019expected], we chose to use only ECCFIC in the name of the package because ECD belongs to the family of generalized ECCFIC measures. However, to preserve the unique implementations of the different methods from the two papers [@yin2019new; @ke2019expected], different R functions are separately created and made available for the ECD measures as well as the ECCFIC measures. As a result, R functions based on ECD are either named "ecd" or prefixed by "ecd", with the ECCFIC related R functions also named in a similar fashion.
 

## Installation

You can install the released version of `eccficm` from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("eccficm")
```

## Available Functions

The main functions are `ecd`, `eccfic`, `ecd.test`, and `eccfic.test` as described below.
 
 - `ecd`, `eccfic`: compute the covariance and correlation type statistics.

 - `ecd.test`, `eccfic.test`: perform a permutation test of independence based on the ECD and ECCFIC correlation and covariance type statistics, respectively.

- `generateData`: generates optional data for feature screening based on
 some specific models.

 - `fscreen`: Performs the sure independence feature screening using
 correlation learning between each predictor and the response.
 
 > Although the slicing method for `ECD` and `ECCFIC` is not very sensitive to the number of slices (ns), it is suggested that the number of data points in each slice should exceed *5* but not close to *n/2*, where *n* is the sample size.

## Examples

### Aircraft data: Speed versus Span

In this example, we use the aircraft data available in the `sm` package in R. The data contains six characteristics of aircraft designs in the twentieth century. We consider two variables, wing span (m) and maximum speed (km/h) in period 3 with 230 observations. Our goal is to test the independence of log(Span) and log(Speed) using ECCFIC correlations as test statistics. Results based on Pearson's product-moment correlatoin and distance correlation are included for comparison. *999* replicates are used for the permutation test. 


We first load the `sm` package to get access to the `aircraft` data and filter the `aircraft` data to obtain data for just the third brand time period.

```{r echo=T, message=F}
library(sm)

# get the aircraft data
data("aircraft", package = "sm")

# get data for the third brand time period
aircraft_p3 <- aircraft[aircraft$Period==3, ] 

attach(aircraft_p3) # expose variable names to the R search path
```

#### Scatter plot of log(Span) against log(Speed)

Create a scatter plot to explore the relationship between `log(Speed)` and `log(Span)`.

```{r Speed-vs-Span, echo=T, message=F}
library(ggplot2)

ggplot(aircraft_p3, aes(log(Span), log(Speed))) +
  geom_point(alpha = 0.8, size=3, shape=1) +
  geom_smooth(method = "loess", se=F) +
  theme_classic()

# plot(log(Speed), log(Span))
```

The above graph suggests the existence of some sort of association between `Speed` and `Span` which appears to be nonlinear. In the subsequent codes we will use the hypothesis testing functions in the `eccficm` package as well as one from the `energy` and `stats` packages to determine if there is enough evidence to conclude on the observed relationship.


At this point we load our software package `eccficm`. 

```{r echo=T}
library(eccficm)
```

#### ECD and ECCFIC association statistics

Let's first take a look at the `eccfic` statistics.
```{r echo=T}
# compute ecd statistics
ecd(log(Speed), log(Span), ns=10)

# compute eccfic statistics
eccfic(log(Speed), log(Span), est = 'slice', ns=10)
```

#### Hypothesis Test of Independence

> The alternative hypothesis of interest is that `log(Speed)` is a function of `log(Span)`. In other words, we want to test whether the `log(Speed)` depends on `log(Span)`.

We first test the independence of `log(Speed)` and `log(Span)` with the `ecd.test` function. We use 10 slices for the slicing estimation, ECD correlation as test statistic, and perform 999 permutations.

```{r echo=T}
ecd.test(log(Speed), log(Span), est="slice", ns=10, ts="ecdcor",  B=999)
```

Using the ECD correlation as test statistic, we obtain a p-value of **0.001**.

Similarly, we use the `eccfic.test` function with 10 slices for its slicing estimation, ECCFIC correlation as test statistic, and 999 total permutations. The p-value for this test is also **0.001**. 

```{r echo=T}
eccfic.test(log(Speed), log(Span), est="slice", ns=10, ts="eccficcor",  B=999)
```


Finally, we compare the results for our independence measures with that of the distance correlation with the same number of permutations as well as the Pearson's product-moment correlation. The DC independence testing function is available in the `energy` package. Interestingly, using DC as test statistic produces the same p-value of `0.001`. However, the p-value associated with the test based on Pearson's correlation is `0.8001`.

```{r echo=T}
# Distance correlation as test statistic
energy::dcor.test(log(Speed), log(Span), R=999)

# Pearson's product-moment as test statistic
stats::cor.test(log(Speed), log(Span))
```


>From the above outputs of test results,  despite the nonlinear relation between wing span and maximum speed, the `ecd.test`, `eccfic.test` and `dcor.test` functions returned  a very small p-value of `0.001`, suggesting that maximum speed of the aircraft depends on its wing span. This shows how sensitive distance correlation and ECCFIC correlations are to nonlinear associations. On the other hand, the Pearson's correlation test suggested no evidence of association, showing how Pearson's correlation cannot detect non-linear associations if one truly exists. Moreover, we see how comparable our measures are to the distance correlation. 

# References{-}
