---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ecdmeasures

<!-- badges: start -->
<!-- badges: end -->

The goal of ecdmeasures is to provide new distance correlation/covariance type of measures for testing independence between two random vectors, which uses the expected difference of conditional and marginal characteristic functions.

## Installation

You can install the released version of ecdmeasures from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("ecdmeasures")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(ecdmeasures)
library(MASS) # for the mvrnorm function
 n=30
 p=5
 q=1
 mu=rep(0,(p+q))
 sigma=diag((p+q))
 DATA=mvrnorm(n, mu, sigma)
 X=DATA[,1:p]
 Y=DATA[,(p+1):(p+q)]
 # using slicing estimation method on a continuous Y
 ecdcor.test(x = X, y = Y, method = "slice", B = "default", ns = 5)
 # using Gaussian kernel estimation method
 ecdcor.test(x = X, y = Y, method = "kernel.gau", B = "default")
 # using Epanechnikov kernel estimation method
 ecdcor.test(x = X, y = Y, method = "kernel.epa", B = "default")

 #This example shows how to directly perform the test if Y is a categorical
 #variable
 library(boot)
 #gravity data is in the boot package,there are 81 measurements in a series
 # of eight experiments
 #gravity$series shows the categories 1-8 of the observations. 999 permutations
 # are used to compute p-value.
str(gravity)
 ecdcor.test(gravity$g, gravity$series, method="slice", B=999)
```


